# Transformation

We know that nuvirdi has an unusually heavy tale so we'll start by transforming our response variable using boxcox.

```{r, fig.cap="Boxcox plot for the model and the model without outliers. No outlier plot is on the right", warning=FALSE}
par(mfrow=c(1,2))
boxcox(lm.all)
boxcox(lm.allNoOutlier)
par(mfrow=c(1,1))
```

```{r}
bcAll <- boxcox(lm.all, plotit = FALSE)
bcNO <- boxcox(lm.allNoOutlier, plotit = FALSE)

# Get the highest value from the boxcox plots
lambdaAll <- with(bcAll, x[which.max(y)])
lambdaNO <- with(bcNO, x[which.max(y)])

# Transform the train and test datasets
trainBCAll <- bcTransF(train, lambdaAll)
testBCAll <- bcTransF(test, lambdaAll)
trainBCNO <- bcTransF(trainNO, lambdaNO)
testBCNO <- bcTransF(test, lambdaNO)

lm.allBc <- lm(nuvirdi ~ ., data = trainBCAll)
lm.allNoOutlierBc <- lm(nuvirdi ~ ., data = trainBCNO)
Radj.allBc <- CalculateRadjLambda(lm.allBc, testBCAll, lambdaAll) # 0.8488961 rétt: 0.7643883
Radj.allNOBc <- CalculateRadjLambda(lm.allNoOutlierBc, testBCNO, lambdaNO)

```

Here we get $R_{adj. full model}=`r Radj.allBc`$ and $R_{adj. NoOutliers}=`r Radj.allNOBc`$ by using the Boxcox method on the response variable. We can see from our ggpairs plot that the explanatory variable $ibm2$ has a heavy tale so lets try to transform that variable as well. Here we use a log transformation for the $ibm2$ variable.

```{r}

# Transform ibm2
trainBCAllIBM2 <- trainBCAll
trainBCAllIBM2$ibm2 <- log(trainBCAll$ibm2)
testBCAllIBM2 <- testBCAll
testBCAllIBM2$ibm2 <- log(testBCAll$ibm2)
trainBCNOIBM2 <- trainBCNO
trainBCNOIBM2$ibm2 <- log(trainBCNO$ibm2)
testBCNOIBM2 <- testBCNO
testBCNOIBM2$ibm2 <- log(testBCNO$ibm2)

lm.allBcIBM2 <- lm(nuvirdi ~ ., data = trainBCAllIBM2)
lm.allNOBcIBM2 <- lm(nuvirdi ~ ., data = trainBCNOIBM2)
Radj.allBcIBM2 <- CalculateRadjLambda(lm.allBcIBM2, testBCAllIBM2, lambdaAll) # 0.8880478 rétt: 0.8342736
Radj.allNOBcIBM2 <- CalculateRadjLambda(lm.allNOBcIBM2, testBCNOIBM2, lambdaNO) 


# Geymi hér. Þetta er besta módelið hingað til
bcBest <- boxcox(lm.best)
lambdaBest <- with(bcBest, x[which.max(y)])
trainBCBest <- bcTransF(train, lambdaBest)
testBCBest <- bcTransF(test, lambdaBest)
trainBCBestIBM2 <- trainBCBest
trainBCBestIBM2$ibm2 <- log(trainBCBest$ibm2)
testBCBestIBM2 <- testBCBest
testBCBestIBM2$ibm2 <- log(testBCBest$ibm2)

lm.best <- lm(nuvirdi ~ ibm2*matssvaedi+undirmatssvaedi+ibteg+teg_eign+fjeld+k.ar, data = trainBCBestIBM2)
Radj.best <- CalculateRadjLambda(lm.best, testBCBestIBM2, lambdaBest)
```

Here we get a better value for R adjusted. We get $R_{adj. full model}=`r Radj.allBcIBM2`$ for the full model and $R_{adj. NoOutliers}=`r Radj.allNOBcIBM2`$ for the model without outliers.

