# Transformation

We know that nuvirdi has an unusually heavy tale so we'll start by transforming our response variable using boxcox.

```{r, echo=FALSE,  fig.cap="Boxcox plot for the four models. Top right: Model with all the training data, top left: Model with no outliers, bottom right: Model with no influential points and bottom left: Model with no outliers and no influential points.", warning=FALSE}
par(mfrow=c(2,2))
boxcox(lm.all)
boxcox(lm.allNoOutlier)
boxcox(lm.allNoInfluential)
boxcox(lm.allNoInflueNoOutlier)
par(mfrow=c(1,1))
```



```{r}
Radj.ALLBC <- BCTranformResponseRadj(lm.all, train, test)
Radj.NOBC <- BCTranformResponseRadj(lm.allNoOutlier, trainNO, test)
Radj.NIBC <- BCTranformResponseRadj(lm.allNoInfluential, trainNoInflu, test)
Radj.NONIBC <- BCTranformResponseRadj(lm.allNoInflueNoOutlier, trainNONI, test)

```

Here below we can see the $R_{adj}$ for the four models after transforming the response variable. $R_{adj}$ is calculated using the test set.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
     & No changes & No outl. & No infl. & No outl. and no infl.  \\ 
  \hline
  $R_{adj}$ & $`r Radj.ALLBC`$ & $`r Radj.NOBC`$ & $`r Radj.NIBC`$ & $`r Radj.NONIBC`$ \\ 
   \hline
\end{tabular}
\end{table}

From the ggpairs image we can see that $ibm2$ has a heavy right tail as well so lets try log-transforming that variable to see if we get better results.


```{r}

Radj.AllBCAndIBM2 <- TransformBCandIBM2(lm.all, train, test)
Radj.NOBCAndIBM2 <- TransformBCandIBM2(lm.allNoOutlier, train, test)
Radj.NIBCAndIBM2 <- TransformBCandIBM2(lm.allNoInfluential, trainNoInflu, test)
Radj.NONIBCAndIBM2 <- TransformBCandIBM2(lm.allNoInflueNoOutlier, trainNONI, test)

```

Here below we can see the $R_{adj}$ for the four models after transforming the response variable and $ibm2$. $R_{adj}$ is calculated using the test set. Now we get much better results for $R_{adj}$.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
     & No changes & No outl. & No infl. & No outl. and no infl.  \\ 
  \hline
  $R_{adj}$ & $`r Radj.AllBCAndIBM2`$ & $`r Radj.NOBCAndIBM2`$ & $`r Radj.NIBCAndIBM2`$ & $`r Radj.NONIBCAndIBM2`$ \\ 
   \hline
\end{tabular}
\end{table}

