
# Variable selection

We saw from the transformation chapter that we got the best models by transforming both the response variable and $ibm2$. So we'll be using those models for variable selection.

```{r, results='hide'}
# Feching models and datasets
ALL <- GetBCandIBM2ModelAndDt(lm.all,train, test)
NO <- GetBCandIBM2ModelAndDt(lm.allNoOutlier, trainNO, test)
NI <- GetBCandIBM2ModelAndDt(lm.allNoInfluential, trainNoInflu, test)
NONI <- GetBCandIBM2ModelAndDt(lm.allNoInflueNoOutlier, trainNONI, test)
```

Lets try to use BIC and AIC criteria to select our variables.

```{r, results='hide'}

# BIC tests
ALLBIC <- findBestBICModel(lm(nuvirdi ~ 1, data = ALL$train), 
                           ALL$model, ALL$train, ALL$test, ALL$lambda)
NOBIC <- findBestBICModel(lm(nuvirdi ~ 1, data = NO$train), 
                          NO$model, NO$train, NO$test, NO$lambda)
NIBIC <- findBestBICModel(lm(nuvirdi ~ 1, data = NI$train), 
                          NI$model, NI$train, NI$test, NI$lambda)
NONIBIC <- findBestBICModel(lm(nuvirdi ~ 1, data = NONI$train), 
                            NONI$model, NONI$train, NONI$test, NONI$lambda)

# AIC tests
ALLAIC <- findBestAICModel(lm(nuvirdi ~ 1, data = ALL$train), 
                           ALL$model, ALL$train, ALL$test, ALL$lambda)
NOAIC <- findBestAICModel(lm(nuvirdi ~ 1, data = NO$train), 
                          NO$model, NO$train, NO$test, NO$lambda)
NIAIC <- findBestAICModel(lm(nuvirdi ~ 1, data = NI$train), 
                          NI$model, NI$train, NI$test, NI$lambda)
NONIAIC <- findBestAICModel(lm(nuvirdi ~ 1, data = NONI$train), 
                            NONI$model, NONI$train, NONI$test, NONI$lambda)

```

We can see that we get the best $R_{adj}$ when using the AIC crite

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
     & No changes & No outl. & No infl. & No outl. and no infl.  \\ 
  \hline
  $R_{adj(BIC)}$ & $`r ALLBIC$Radj`$ & $`r NOBIC$Radj`$ & $`r NIBIC$Radj`$ & $`r NONIBIC$Radj`$ \\ 
  \hline
  $R_{adj(AIC)}$ & $`r ALLAIC$Radj`$ & $`r NOAIC$Radj`$ & $`r NIAIC$Radj`$ & $`r NONIAIC$Radj`$ \\ 
   \hline
\end{tabular}
\end{table}

Lets now try something different. Lets use the transformed data without any changes and use the add1 function to add explanatory variables.

```{r, results='hide'}
add1(lm(nuvirdi~1, data = ALL$train),~ ibm2 + kdagur + matssvaedi + teg_eign + undirmatssvaedi +
       haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+lyfta+
       fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
```

Lets start by adding ibm2.

```{r, results='hide'}
add1(lm(nuvirdi~ibm2, data = ALL$train),~ ibm2 + kdagur + matssvaedi + teg_eign + undirmatssvaedi + 
       haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+lyfta+
       fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
```

Lets now add matssvaedi. First lets see what model to use.

```{r}
drop1(lm(nuvirdi~ibm2*matssvaedi, data = ALL$train), test = "F")
```

We can see from the drop1 function that the best model seems to have different slope and different intercept when just using ibm2 and matssvaedi. Lets continue adding variables.

```{r, results='hide'}
lm.temp <- lm(nuvirdi~ibm2*matssvaedi, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+   fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+lyfta+  
fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Adding kdagur
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+   fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+  
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Adding teg_eign
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
       lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Adding byggar
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Adding undirmatssvaedi
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
       lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Adding haednr
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
       lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Adding fjhaed
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Adding fjbilast
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed+fjbilast, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Adding fjstof
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed+fjbilast+
                fjstof, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr+ fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# adding lyfta
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed+fjbilast+fjstof+
                lyfta, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.lyfta <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.add1lyftaTrain <- summary(lm.temp)$adj.r.squared
# adding fjsturt
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed+fjbilast+fjstof+
                lyfta+fjsturt, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.fjsturt <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.add1fjsturtTrain <- summary(lm.temp)$adj.r.squared
# adding stig10
lm.temp <- lm(nuvirdi~ibm2*matssvaedi+kdagur+teg_eign+byggar+
                undirmatssvaedi+haednr+fjhaed+fjbilast+fjstof+
                lyfta+fjsturt+stig10, data = ALL$train)
add1(lm.temp,~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+ibteg+k.ar+byggar+
lyfta+fjklos+fjeld+fjherb ,data=ALL$train, test = "F")
Radj.add1Final <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.add1FinalTrain <- summary(lm.temp)$adj.r.squared
```

The table below shows $R_{adj}$ for the last three steps when using the add1 function.

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
     & Add lyfta & Add fjsturt & Add stig10  \\ 
  \hline
  $R_{adj(add1)Test}$ & $`r Radj.lyfta`$ & $`r Radj.fjsturt`$ & $`r Radj.add1Final`$ \\ 
  \hline
  $R_{adj(add1)}Train$ & $`r Radj.add1lyftaTrain`$ & $`r Radj.add1fjsturtTrain`$ & $`r Radj.add1FinalTrain`$ \\ 
   \hline
\end{tabular}
\end{table}

After using the add1 function until there was no significant explanatory variable left we got $R_{adj}=`r Radj.add1Final`$. 
Lets try using drop1 instead with different intercept and slope for matsvaedi.

```{r, results='hide'}
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                ibteg+k.ar+byggar+lyfta+fjklos+fjeld+fjherb, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Dropping k.ar
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                ibteg+byggar+lyfta+fjklos+fjeld+fjherb, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda)
# Dropping fjklos
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                ibteg+byggar+lyfta+fjeld+fjherb, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Dropping fjeld
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                ibteg+byggar+lyfta+fjherb, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.temp <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
# Dropping ibteg
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi +
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                byggar+lyfta+fjherb, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.drIbteg <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.drIbtegTrain <- summary(lm.temp)$adj.r.squared
# Dropping fjherb
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
                byggar+lyfta, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.drfjherb <- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.drfjherbTrain <- summary(lm.temp)$adj.r.squared
# Dropping fjbkar
lm.temp <- lm(nuvirdi ~ ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
                haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+byggar+
                lyfta, data = ALL$train)
drop1(lm.temp, test = "F")
Radj.drfjbkar<- CalculateRadjLambda(lm.temp, ALL$test, ALL$lambda) 
Radj.drfjbkarTrain <- summary(lm.temp)$adj.r.squared
```

Table below shows $R_{adj}$ for the last three steps when using the drop1 function.

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
     & Drop ibteg & Drop fjherb & Drop fjbkar  \\ 
  \hline
  $R_{adj(drop1)Test}$ & $`r Radj.drIbteg`$ & $`r Radj.drfjherb`$ & $`r Radj.drfjbkar`$ \\ 
  \hline
  $R_{adj(drop1)Train}$ & $`r Radj.drIbtegTrain`$ & $`r Radj.drfjherbTrain`$ & $`r Radj.drfjbkarTrain`$ \\ 
   \hline
\end{tabular}
\end{table}

Lets now try to use BIC and AIC starting with the model (nuvirdi ~ ibm2*matssvaedi ).

```{r, results='hide'}
null <- lm(nuvirdi~ibm2*matssvaedi, data = ALL$train)
full <- lm(nuvirdi~ibm2*matssvaedi + kdagur + teg_eign + undirmatssvaedi + 
             haednr + fjhaed+fjstof+byggar+fjsturt+stig10+fjbilast+fjbkar+
             ibteg+k.ar+byggar+lyfta+fjklos+fjeld+fjherb, data = ALL$train)
# BIC tests
ALLBIC <- findBestBICModel(null, full, ALL$train, ALL$test, ALL$lambda)
RadjBICTrain <- summary(ALLBIC$model)$adj.r.squared
# AIC tests
ALLAIC <- findBestAICModel(null, full, ALL$train, ALL$test, ALL$lambda)
RadjAICTrain <- summary(ALLAIC$model)$adj.r.squared
```

The table below shows the best $R_{adj}$ for each test when we start with different intercepts for matssvaedi.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
     & add1 & drop1 & AIC & BIC  \\ 
  \hline
  $R_{adjTest}$ & $`r Radj.fjsturt`$ & $`r Radj.drfjbkar`$ & $`r ALLAIC$Radj`$ & $`r ALLBIC$Radj`$ \\ 
  \hline
  $R_{adjTrain}$ & $`r Radj.add1fjsturtTrain`$ & $`r Radj.drfjbkarTrain`$ & $`r RadjAICTrain`$ & $`r RadjBICTrain`$ \\ 
   \hline
\end{tabular}
\end{table}
